---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---




## Loading and preprocessing the data


```{r load , echo=TRUE}
    setwd("C:/Users/chaudharyan/Desktop/RepData_PeerAssessment1")
    file<-read.csv("./activity/activity.csv", header=TRUE)
    
     
```
## What is mean total number of steps taken per day?
    
```{r meansteps, echo=TRUE}
    data<-na.omit(file)
    s<-aggregate(list(Steps=data$steps), list(Date=data$date), sum)
    hist(s$Steps, main="Total number of steps taken each day", xlab="Steps",col="grey")
    mean(s$Steps)
    median(s$Steps)
```

## What is the average daily activity pattern?
    
A time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days
```{r average , echo=TRUE}   
data<-na.omit(file)   
m<-aggregate(list(MeanSteps=data$steps), list(Interval=data$interval), mean)
plot(m$Interval, m$MeanSteps, type="l", xlab="Interval", ylab="Mean Steps",main="Average number of steps")

```


The 5-minute interval, on average across all the days in the dataset, which contains the maximum number of steps

```{r maxaverage, echo=TRUE}
m[which.max(m$MeanSteps),]

```


## Imputing missing values

Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r include=FALSE}
library(dplyr)
```

```{r missing, echo=TRUE}

    
    na_rows<-nrow(filter(file, is.na(file$steps)==TRUE))
    
```

Calcualting mean of every 5-minute interval and creating new dataset with missing value filled in with the mean

```{r fillthemissing, echo=TRUE}  

    o <- na.omit(file) 
    Int_mean <- aggregate(list(Steps=o$steps), list(Interval=o$interval), mean)
    
    
    f<-file

    for(i in seq(0,2355,5)) 
    {
        f[f$interval==i & is.na(f$steps)==TRUE,1]<- Int_mean[Int_mean$Interval==i,2]
        
    }
    h<-aggregate(list(Steps=f$steps), list(Date=f$date), sum)
    hist(h$Steps, main="Total number of steps taken each day", xlab="Steps",col="grey")
    mean(h$Steps)
    median(h$Steps)
    
```

When I compare the mean and median value calcualted above with the first part of asingnment, I find no difference in mean value and a very small difference in median value . 


First Part of assignment 
```{r}
    mean(s$Steps)
    median(s$Steps)
```

Current Assignment 
```{r}
    mean(h$Steps)
    median(h$Steps)
```

When I compare the the total steps from then first part of the assignment with the 5-minute interval filled NA values, I find an increase of 86,129 in the total numbers. This implies that filling the missing values does impact the total daily number of steps.

First Part of assignment 
```{r}
    sum(s$Steps)
```

Current Assignment 
```{r}
    sum(h$Steps)
```



## Are there differences in activity patterns between weekdays and weekends?

```{r weekday, echo=TRUE}
f$day<-weekdays(as.Date(f$date))
f$Week<-factor(f$day %in% c("Monday","Tuesday","Wednesday","Thursday","Friday"),levels=c(TRUE,FALSE),         labels=c("weekday","weekend") )

ag_Day<-aggregate(list(Steps=f$steps), list(Wkday=f$Week, Interval=f$interval), mean)

library(ggplot2)
ggplot(ag_Day,aes(x=Interval, y=Steps))+geom_line(color="dark blue")+facet_wrap(~Wkday) +labs(x="Interval", y="Number of steps")


```